---
title: "Mini-Project 2"
authors: "Marisa Youngblood, Tara Marin, and Jordan Moody"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
load("individuals.rda")
```

##
We want to see what trends or patterns have emerged since the Citizens United case was decided in 2010. In particular, what has happened to political contributions from for-profit corporations since the decision. If there has been a notable increase, has a particular party benefitted disproportionally, how has the money been channeled to campaigns – individual candidates, PACs, etc since 2010?

```{r}
```

##
In the most recently presidential election, both Bernie Sanders and Donald Trump's campaigns received a surge of support from individuals donating small amounts rather than corporate donors and super PACs, demonstrating grassroots appeal. We want to see what % of their funding came from small donors and compare that to the % that came from other sources using a bar graph. 

```{r}
small_individual <- individuals %>%
  filter(transaction_amt >= 1, transaction_amt <= 200)

#Filtering for small individual donations under $200

small_committees <- committees %>%
  inner_join(small_individual, by = "cmte_id")

#Using inner_join function to connect individual donations to their corresponding committee 
```

###Grading Rubric
- Using at least two of the five basic verbs
- Annotating
- Writing a function
- Using at least one join
- Blog post text
- Using GitHub
- Issues tab